---
title: "Assignment_1_Beissner"
author: "Ben Ole Beissner"
date: "2023-03-03"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
# Load necessary libraries and data
library(readr)
library(ggplot2)
library(tidyr)
library(MASS)
library(tibble)
library(dagitty)
library(ggdag)
a1_data <- data.frame(read_csv("~/Computational Statistics/Assignment 1/a1_data.csv"))
#a1_data
```
```{r}
# Histogram plot of variables
ggplot(pivot_longer(a1_data, everything()), aes(value)) +
  geom_histogram(bins = 100) +
  facet_wrap(~name, nrow = 3)
```
```{r}
# Mean of each variable
colMeans(a1_data)
```
```{r}
# Variance of each variable
sapply(a1_data, var)
```


# C) Implementation in Python
A Pearson’s correlation test is ran to calculate the partial correlation between all variables, in case of no correlation
the third DAG is printed. Here we assume that only partial correlations with a value greater than 0.15 or in the negative case less than -0.15 are significantly correlated. Note that with a lower treshold, more linear relationships between variables are taken into account which leads to different DAGs.


```{r}
# Computing partial correlation between all variables
#install.packages("ppcor")
library(ppcor)
partial_corr = pcor(a1_data)
pcor_est = data.frame(partial_corr$estimate)


# filter for pairs of variables with significant partial correlation (> 0.15) - comment out to see all partial correlations between variables
filter_df <- function(pcor_est) {
  # Define the condition to filter by
  condition <- pcor_est > 0.15
  
  # Apply the condition to the entire dataframe
  filtered_df <- pcor_est * condition
  # Return the filtered dataframe
  print(filtered_df)
  return(filtered_df)
}
filter_df(pcor_est)
```
```{r}
# filter for pairs of variables with significant partial correlation (> 0.15) 

filter_df <- function(pcor_est) {
  # Define the condition to filter by
  condition <- pcor_est < -0.15
  
  # Apply the condition to the entire dataframe
  filtered_df <- pcor_est * condition
  # Return the filtered dataframe
  print(filtered_df)
  return(filtered_df)
}
filter_df(pcor_est)
```


Because of our assumption of normal distributed variables we can use a´the F-test to test the variances. We use the null hypothesis that the ration between Var(x) and Var(Y) is 1, alternative it is less than 1. According to which variance is bigger, we print the DAG. We do this for all partial correlated pairs of variables seperately and afterwards conclude the bivariate relationships into one DAG:

```{r}
# F-test for all variables with significant correlation to determine the direction of the arrow
#A and B
res.ftest <- var.test(a1_data$A, a1_data$B, data=a1_data, alternative = "less")
res.ftest
```

```{r}
# Print bivariate DAG according to F-test
if (res.ftest$estimate< 1){
  dag <- dagitty("dag{B <- A}")
  tidy_ggdag<-tidy_dagitty(dag)
}else {
  dag <- dagitty("dag{A <- B}")
  tidy_ggdag<-tidy_dagitty(dag)
}
ggdag(tidy_ggdag) +
  theme_dag()
```

```{r}
#F and G
res.ftest <- var.test(a1_data$F, a1_data$G, data=a1_data, alternative = "less")
res.ftest
```
```{r}
#DAG F and G
if (res.ftest$estimate< 1){
  dag <- dagitty("dag{G <- F}")
  tidy_ggdag<-tidy_dagitty(dag)
}else {
  dag <- dagitty("dag{F <- G}")
  tidy_ggdag<-tidy_dagitty(dag)
}
ggdag(tidy_ggdag) +
  theme_dag()
```

```{r}
#I and J
res.ftest <- var.test(a1_data$I, a1_data$J, data=a1_data, alternative = "less")
res.ftest
```
```{r}
#DAG I and J
if (res.ftest$estimate< 1){
  dag <- dagitty("dag{J <- I}")
  tidy_ggdag<-tidy_dagitty(dag)
}else {
  dag <- dagitty("dag{I <- J}")
  tidy_ggdag<-tidy_dagitty(dag)
}
ggdag(tidy_ggdag) +
  theme_dag()
```

```{r}
#D and E
res.ftest <- var.test(a1_data$D, a1_data$E, data=a1_data, alternative = "less")
res.ftest
```
```{r}
if (res.ftest$estimate< 1){
  dag <- dagitty("dag{E <- D}")
  tidy_ggdag<-tidy_dagitty(dag)
}else {
  dag <- dagitty("dag{D <- E}")
  tidy_ggdag<-tidy_dagitty(dag)
}
ggdag(tidy_ggdag) +
  theme_dag()
```

```{r}
#F and H
res.ftest <- var.test(a1_data$F, a1_data$H, data=a1_data, alternative = "less")
res.ftest
```
```{r}
if (res.ftest$estimate< 1){
  dag <- dagitty("dag{H <- F}")
  tidy_ggdag<-tidy_dagitty(dag)
}else {
  dag <- dagitty("dag{F <- H}")
  tidy_ggdag<-tidy_dagitty(dag)
}
ggdag(tidy_ggdag) +
  theme_dag()
```

```{r}
#F and J
res.ftest <- var.test(a1_data$F, a1_data$J, data=a1_data, alternative = "less")
res.ftest
```
```{r}
if (res.ftest$estimate< 1){
  dag <- dagitty("dag{J <- F}")
  tidy_ggdag<-tidy_dagitty(dag)
}else {
  dag <- dagitty("dag{F <- J}")
  tidy_ggdag<-tidy_dagitty(dag)
}
ggdag(tidy_ggdag) +
  theme_dag()
```

```{r}
#G and H
res.ftest <- var.test(a1_data$G, a1_data$H, data=a1_data, alternative = "less")
res.ftest
```
```{r}
if (res.ftest$estimate< 1){
  dag <- dagitty("dag{H <- G}")
  tidy_ggdag<-tidy_dagitty(dag)
}else {
  dag <- dagitty("dag{G <- H}")
  tidy_ggdag<-tidy_dagitty(dag)
}
ggdag(tidy_ggdag) +
  theme_dag()
```

```{r}
#I and K
res.ftest <- var.test(a1_data$I, a1_data$K, data=a1_data, alternative = "less")
res.ftest
```
```{r}
if (res.ftest$estimate< 1){
  dag <- dagitty("dag{K <- I}")
  tidy_ggdag<-tidy_dagitty(dag)
}else {
  dag <- dagitty("dag{I <- K}")
  tidy_ggdag<-tidy_dagitty(dag)
}
ggdag(tidy_ggdag) +
  theme_dag()
```

```{r}
#J and B
res.ftest <- var.test(a1_data$J, a1_data$B, data=a1_data, alternative = "less")
res.ftest
```
```{r}
if (res.ftest$estimate< 1){
  dag <- dagitty("dag{B <- J}")
  tidy_ggdag<-tidy_dagitty(dag)
}else {
  dag <- dagitty("dag{J <- B}")
  tidy_ggdag<-tidy_dagitty(dag)
}
ggdag(tidy_ggdag) +
  theme_dag()
```
```{r}
#I and H
res.ftest <- var.test(a1_data$I, a1_data$H, data=a1_data, alternative = "less")
res.ftest
```
```{r}
if (res.ftest$estimate< 1){
  dag <- dagitty("dag{H <- I}")
  tidy_ggdag<-tidy_dagitty(dag)
}else {
  dag <- dagitty("dag{I <- H}")
  tidy_ggdag<-tidy_dagitty(dag)
}
ggdag(tidy_ggdag) +
  theme_dag()
```

```{r}
#J and A
res.ftest <- var.test(a1_data$J, a1_data$A, data=a1_data, alternative = "less")
res.ftest
```
```{r}
if (res.ftest$estimate< 1){
  dag <- dagitty("dag{A <- J}")
  tidy_ggdag<-tidy_dagitty(dag)
}else {
  dag <- dagitty("dag{J <- A}")
  tidy_ggdag<-tidy_dagitty(dag)
}
ggdag(tidy_ggdag) +
  theme_dag()
```

```{r}
#J and H
res.ftest <- var.test(a1_data$J, a1_data$H, data=a1_data, alternative = "less")
res.ftest
```
```{r}
if (res.ftest$estimate< 1){
  dag <- dagitty("dag{H <- J}")
  tidy_ggdag<-tidy_dagitty(dag)
}else {
  dag <- dagitty("dag{J <- H}")
  tidy_ggdag<-tidy_dagitty(dag)
}
ggdag(tidy_ggdag) +
  theme_dag()
```

```{r}
# Print whole DAG with partial correlation
g <- dagitty('dag{
  A [pos="-2,2"]
  B [pos="-3,1"]
  C [pos="-2,1"]
  D [pos="-2,0"]
  E [pos="-3,-1"]
  F [pos="2,0"]
  G [pos="4,0"]
  H [pos="0,-1"]
  I [pos="2,3"]
  J [pos="0,2"]
  K [pos="2,1"]
  
  B -> A -> J
  B -> J
  E -> D
  G -> F
  H -> I -> J
  H -> J
  I -> K
}')
plot(g)
```